Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2018-10-21 paper/paper.tex
Change exp part

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of porous graphene (PG). Two
reservoirs, namely the high-concentration reservoir (HCR) containing
electrolyte solution with molar concentration $c_0$, and the
low-concentration reservoir (LCR) containing de-ionized (DI) water,
are separated by a PG membrane supported by polycarbonate track-etched
(PCTE) film. Magnetic stirrers are used to minimize the effects of
external concentration polarization, and a conductivity probe is
placed in the LCR to monitor the ionic concentration as a function of
time. The membrane is attached to a piece of copper tape connected to
a voltage source applying an electrical bias $V_{\mathrm{G}}$, with
the LCR grounded (\Fig{} \ref{fig:1}a inset). We fabricated the
PCTE-supported PG using the protocol developed by our group
\cite{Choi_2018}, as schematically shown in \Fig{} \ref{fig:1}b. The
normally distributed, circular pores with pore diameters of 20$\pm$10
nm (\Fig{} \ref{fig:1}c inset) over a large area were drilled on
graphene.  By optimizing the subsequent transfer process, a high
surface coverage ($>$ 98\%) of graphene on PCTE was reached (\Fig{}
\ref{fig:1}c), allowing to reliably characterize the ion transport
behavior (details see Supplementary Information Section S1).  The
system presented here allows us to systematically investigate the
diffusive flux across a sheet of porous graphene, $\boldsymbol{J}$,
driven by the gradient of electrochemical potential in the electrolyte
medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity versus time profile
(e.g. \Fig{} \ref{fig:2}a) remains linear within the measurement
duration (1 h), indicating the cross-membrane concentration gradient
is almost constant during the experiment. The diffusive fluxes
correspond to different salts were obtained by converting the
conductivity-time profiles with the calibration curves (blue bars in
\Fig{} \ref{fig:2}b and Supplementary \Fig{} S2). For comparison
purposes, based on the assumption that the membrane consists
cylindrical pores with uniform diameter and a low tortuosity $\tau$ of
$\sim{}$1.2 \cite{O_Hern_2012}, the molar diffusive flux through the
bare PCTE membrane, $\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

@@

corresponds to the species (anion or cation), $D$ is the diffusivity,

@@

where $x$ is the molar fraction, $z$ is the ionic valence, $e$ is the
unit charge, and $\psi$ is the electric potential. On the other hand,
the Poisson equation describing the electric potential distribution is
given by:

@@

individual materials in the system. Including water and the internal

@@

Eq. \ref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing Eqs. \ref{eq:pnp}-\ref{eq:poisson}

@@

Eqs. \ref{eq:electro-neutral}-\ref{eq:delta-phiG} that were solved

@@

graphene (Eq. \ref{eq:delta-phiG}). However, for graphene synthesized

@@

\subsection{Salt rejection mechanism}
\label{sec:mechanism}
The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near to a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \ref{eq:pnp} and \ref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye screening length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $R_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes
small. (also see Supplementary \Fig{}s S4). We also
observe that $\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
S5). The observations confirm the distinct ionic
transport pathways upon applying a positive $V_{\mathrm{G}}$: pore
center for cations and pore edge for anions. \Fig{} \ref{fig:4}c
presents the calculated z-component of the cation and anion fluxes
through the pore at the graphene plane (z = 0), $\boldsymbol{J}_{z}$
,as a function of the normalized radius, $r/R_{\mathrm{G}}$ , where
$r$ is the radial coordinate and $R_{\mathrm{G}}$ is the pore radius,
at different $V_{\mathrm{G}}$ values. Note that $\boldsymbol{J}_{z}$
is negative because the LCR is placed at z < 0 in the simulations. At
$V_{\mathrm{G}}$ = 0 (pure diffusion), the anion and cation flux
profiles are identical, with a higher flux near the pore edge, which
is expected, as the concentration gradient is higher. By gradually
increasing $V_{\mathrm{G}}$, both fluxes are reduced throughout the
pore, while the degree of reduction is more pronounced at the pore
edge for the cation, and at the pore center for the anion, following
the scenarios we discussed above. Accordingly, the integrated fluxes
across the pore, $|\boldsymbol{J}_{z}|$, are reduced with
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:4}d). A total flux reduction of up
to 80\% is predicted. Another interesting finding here is that, by
increasing $V_{\mathrm{G}}$, the nanopore transport preferentially
allows cations over anions, or in other words, the anion flux is more
reduced with $V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the
ion selectivity of graphene nanopore \cite{Rollings_2016}.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (Supplementary Section 6). Intriguingly,
by increasing $\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux
starts to increase, reversing the trend observed at the low
$\psi_{\mathrm{G}}$ regime, due to (i) the drift flux become dominant
over diffusion flux and (ii) the concentration near the pore edge
(prefered path) greatly increases. This level of $\psi_{\mathrm{G}}$
cannot be reached experimentally in graphene membrane, as this
requires a $V_{\mathrm{G}}$ larger than 2.0 V, which triggers the
electrochemical reactions \cite{Toh_2011}. On the other hand, in an
IFET, the electric potential on the pore wall is considerably higher,
equivalent to $V_{\mathrm{G}}$ due to the metallic nature of gate
electrode (having an infinitely-large quantum capacitance). We
therefore conclude that the quantum capacitance-induced nonlinear
damping in the surface potential results in the observed salt
rejection.  Following the above discussions, we further investigate
the physical limits for the biased graphene nanopores. Specifically,
the salt rejection characteristics are controlled by: (i) the overlap
of EDL inside the nanopore, essentially controlled by two length
scales of $\lambda_{\mathrm{D}}$ and $R_{\mathrm{G}}$, and (ii) the
graphene surface potential $\psi_{\mathrm{G}}$ controlled by
$V_{\mathrm{G}}$ following Eq. \ref{eq:Vg}. To this end, we calculate
$\psi_{\mathrm{G}}$ as a function of
$\lambda_{\mathrm{D}} / R_{\mathrm{G}}$ and $V_{\mathrm{G}}$ (\Fig{}
\ref{fig:5}a). Clearly, when the bulk concentration $c_{0}$ increases,
a higher $V_{\mathrm{G}}$ is required to reach the same level of
$\psi_{\mathrm{G}}$. Consequently, as revealed in \Fig{} \ref{fig:5}b,
the calculated salt rejection factor $\xi$ increases with both
$\lambda_{\mathrm{D}} / R_{\mathrm{G}}$ and $V_{\mathrm{G}}$, in line
with the experimentally observed $\xi - c_{0}$ dependence. We also
find the overall trends for $\psi_{\mathrm{G}}$ and $\xi$ with respect
to $\lambda_{\mathrm{D}}/R_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / R_{\mathrm{G}}$
= 1.0, a high value of $\xi$ of $>$ 0.85 can be achieved.

>>>>

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm (\Fig{} \ref{fig:1}c inset) over a
large area were drilled on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of graphene on
PCTE was reached (\Fig{} \ref{fig:1}c), allowing to reliably
characterize the ion transport behavior (details see Supplementary
Information Section S1).  The system presented here allows us to
systematically investigate the diffusive flux across a sheet of porous
graphene, $\boldsymbol{J}$, driven by the gradient of electrochemical
potential in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity versus time profile
(e.g. \Fig{} \ref{fig:2}a) remains linear within the measurement
duration (1 h), suggesting the concentration gradient across the
membrane remains constant. The diffusive fluxes
correspond to different salts were obtained by converting the
conductivity-time profiles with the calibration curves (blue bars in
\Fig{} \ref{fig:2}b and Supplementary \Fig{} S2). For comparison
purposes, based on the assumption that the membrane consists
cylindrical pores with uniform diameter and a low tortuosity $\tau$ of
$\sim{}$1.2 \cite{O_Hern_2012}, the molar diffusive flux through the
bare PCTE membrane, $\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

@@

corresponds to the ionic species (anion or cation), $D$ is the diffusivity,

@@

where $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:

@@

individual materials in the system, including water and the internal

@@

Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing Eqs. \eqref{eq:pnp}-\eqref{eq:poisson}

@@

Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved

@@

graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene synthesized

@@

\subsection{Salt rejection mechanism}
\label{sec:mechanism}
The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near to a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \eqref{eq:pnp} and \eqref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye screening length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $R_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes
small. (also see Supplementary \Fig{} S4). We also
observe that $\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
S5). The observations confirm the distinct ionic
transport pathways upon applying a positive $V_{\mathrm{G}}$: pore
center for cations and pore edge for anions. \Fig{} \ref{fig:4}c
presents the calculated \textit{z}-component of the cation and anion fluxes
through the pore at the graphene plane (z = 0), $\boldsymbol{J}_{z}$
,as a function of the normalized radius, $r/R_{\mathrm{G}}$ , where
$r$ is the radial coordinate and $R_{\mathrm{G}}$ is the pore radius,
at different $V_{\mathrm{G}}$ values. Note that $\boldsymbol{J}_{z}$
is negative because the LCR is placed at z$<$0 in the simulation box. At
$V_{\mathrm{G}}$ = 0 (pure diffusion), the anion and cation flux
profiles are identical, with a higher flux near the pore edge, which
is expected, as the concentration gradient is higher. By gradually
increasing $V_{\mathrm{G}}$, both fluxes are reduced throughout the
pore, while the degree of reduction is more pronounced at the pore
edge for the cation, and at the pore center for the anion, following
the scenarios we discussed above. Accordingly, the integrated fluxes
across the pore, $|\boldsymbol{J}_{z}|$, are reduced with
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:4}d). A total flux reduction of up
to 80\% is predicted. Another interesting finding here is that, by
increasing $V_{\mathrm{G}}$, the nanopore transport preferentially
allows cations over anions, or in other words, the anion flux is more
reduced with $V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the
ion selectivity of graphene nanopore \cite{Rollings_2016}.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (Supplementary Section 6). Intriguingly,
by increasing $\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux
starts to increase, reversing the trend observed at the low
$\psi_{\mathrm{G}}$ regime, due to (i) the drift flux become dominant
over diffusion flux and (ii) the concentration near the pore edge
(prefered path) greatly increases (details see Supplementary Section
SXX \todo{markr}). This level of $\psi_{\mathrm{G}}$ cannot be reached
experimentally in graphene membrane, as this requires a
$V_{\mathrm{G}}$ larger than 2.0 V, which triggers the electrochemical
reactions \cite{Toh_2011}. On the other hand, in an IFET, the electric
potential on the pore wall is considerably higher, equivalent to
$V_{\mathrm{G}}$ due to the metallic nature of gate electrode (having
an infinitely-large quantum capacitance). We therefore conclude that
the quantum capacitance-induced nonlinear damping in the surface
potential results in the observed salt rejection.  Following the above
discussions, we further investigate the physical limits for the biased
graphene nanopores. Specifically, the salt rejection characteristics
are controlled by: (i) the overlap of EDL inside the nanopore,
essentially controlled by two length scales of $\lambda_{\mathrm{D}}$
and $R_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / R_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / R_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/R_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / R_{\mathrm{G}}$
= 1.0, a high value of $\xi$ of $>$ 0.85 can be achieved.
